\textbf{Effective application of machine learning to a microbiome-based classification problem}

### 2020-01-22 submission/manuscript.Rmd
updates for rebuttal

We also compared the top 20 OTUs selected by permutation importance in L2-regularized logistic regression and the highest performing tree-based models, random forest and XGBoost. 2 and 5 OTUs, respectively, were shared among the models. These results indicate that we were able to identify a few more OTUs that are deemed important by both linear and non-linear models when we used permutation importance as our interpretation method.

Figure S7:

@@

**Figure S8. Interpretation of the linear ML models with permutation importance.** (A) L1-regularized SVM with linear kernel, (B) L2-regularized SVM with linear kernel, and (C) L2-regularized logistic regression were interpreted using permutation importance using held-out test set.

>>>>

We also compared the top 20 OTUs selected by permutation importance in L2-regularized logistic regression and the highest performing tree-based models, random forest and XGBoost. 2 and 5 OTUs, respectively, were shared among the models. These results indicate that we were able to identify important OTUs that are shared across the highest performing linear and non-linear models when we use permutation importance as our interpretation method.

We then evaluated the difference in relative abundances of the top 20 OTUs identified in L2-regularized logistic regression and random forest models between healthy patients and patients with SRNs [Figure S8]. There were minimal differences in relative abundances across OTUs between different diagnoses. This supports our claim that it is not possible to differentiate disease vs healthy states by focusing on individual taxa.

@@

\newpage
\includegraphics[height=20cm, width=13cm]{Figure_S8.pdf}

**Figure S8. Relative abundances of the 20 most important OTUs in L2-regularized logistic regression and random forest models.** The most important 20 OTUs were chosen for (A) Random forest and (B) L2-regularized logistic regression models by permutation importance and ranking feature coefficients, respectively. The relative abundances of these OTUs were compared based on the diagnosis of the patients. The minimal differences betweeen relative abundances for these OTUs show that it is not possible to differentiate disease vs healthy states by focusing on individual taxa.

### 2020-01-22 submission/manuscript.tex
updates for rebuttal

We also compared the top 20 OTUs selected by permutation importance in
L2-regularized logistic regression and the highest performing tree-based
models, random forest and XGBoost. 2 and 5 OTUs, respectively, were
shared among the models. These results indicate that we were able to
identify a few more OTUs that are deemed important by both linear and
non-linear models when we used permutation importance as our
interpretation method.

Figure S7:

@@

\textbf{Figure S5. NMDS ordination with Bray-Curtis distances.}

@@

\textbf{Figure S6. Permutation importance analysis.} Permutation
importance analysis measures the decrease in the predictive performance
of the model after we permute (A) a feature's or (B) a group of
correlated features' values, which breaks the relationship between the
feature and the actual outcome.

@@

>>>>

We also compared the top 20 OTUs selected by permutation importance in
L2-regularized logistic regression and the highest performing tree-based
models, random forest and XGBoost. 2 and 5 OTUs, respectively, were
shared among the models. These results indicate that we were able to
identify important OTUs that are shared across the highest performing
linear and non-linear models when we use permutation importance as our
interpretation method.

We then evaluated the difference in relative abundances of the top 20
OTUs identified in L2-regularized logistic regression and random forest
models between healthy patients and patients with SRNs {[}Figure S8{]}.
There were minimal differences in relative abundances across OTUs
between different diagnoses. This supports our claim that it is not
possible to differentiate disease vs healthy states by focusing on
individual taxa.

@@

\textbf{Figure S5. NMDS ordination of Bray-Curtis distances.} NMDS
ordination relating the community structures of the fecal microbiota
from 490 patients (261 patients with normal colonoscopies and 229
patients who have screen relevant neoplasias; SRNs).

@@

\textbf{Figure S6. Permutation importance analysis.} Permutation
importance analysis measures the decrease in the predictive performance
of the model after we permute (A) a feature's or (B) a group of
correlated features' values, which breaks the relationship between the
feature and the diagnosis.

@@

\includegraphics[height=20cm, width=13cm]{Figure_S8.pdf}

\textbf{Figure S8. Relative abundances of the 20 most important OTUs in
L2-regularized logistic regression and random forest models.} The most
important 20 OTUs were chosen for (A) Random forest and (B)
L2-regularized logistic regression models by permutation importance and
ranking feature coefficients, respectively. The relative abundances of
these OTUs were compared based on the diagnosis of the patients. The
minimal differences betweeen relative abundances for these OTUs show
that it is not possible to differentiate disease vs healthy states by
focusing on individual taxa.
